### 2010-05-04 doc/workstatement.tex
Big update to use underlying mongodb; added mapping db; modified all services; simplified results generation (will be done in mongodb).
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@392 4525493e-7705-40b1-a816-d608a930855b

\documentclass[a4paper]{jpconf}
\usepackage{graphicx}
\begin{document}
\title{The CMS Data Aggregation System}

\author{Valentin Kuznetsov}
\address{Cornell University, Ithaca, New York, USA}
\ead{vkuznet@gmail.com}

\begin{abstract}
A metadata plays significant role in a large modern enterprises, research experiments,
digital libraries where it comes from different sources and distributed in a 
variety of forms and digital formats. It is organized and managed by constantly
evolving software using both relational and nonrelational data sources. There is
a big demand almost in any industry to access information from multiple 
heterogeneous sources.

The CMS experiment at CERN, represents a particular example where data discovery
plays significant role in daily acitivities of physycists. 
It is expected that starting 2009 it will collect around a few PB of data every 
year of its operations. This reflects into
several hundren of TB of metadata produced by experiment. 
We propose to design and implement a new data aggregation system to search for
data from different relational and non-relational data sources within CMS
experiment environment.
\end{abstract}

\newpage

\section{Introduction}
The CMS experiment is one of the major experiments at CERN. It is expected
that it will start collecting its data in late 2009. Apart from data management
issue the data discovery will play significant role in any analysis and
production workflow tools. In CMS many databases, both
proprietary (ORACLE) and open-source (MySQL, Postgres, SQLite), are used
to store different sort of meta-data. This includes data bookkeeping information,
data location information, run summary information, data quality information,
detector conditions, etc. The information is accessed in a form of
various data-services and presented to the users in variety of data-formats.
Such heterogeneous and distributed environment represents not only manageable
issue, but also inefficient and cumbersome data look-up capabilities. It is desired
to look-up information from more then one data-service and have uniform data
representation across multiple data-services. As a goal we are targeting
towards the Query Language which can be used to place queries across
one or more data-services. Its benefits can be adopted to data-discovery and
production workflow tools.
%An efficient search over one or
%multiple databases without explosure of underlying schema is a key 
%ingredient of any activity in CMS, from production workflow tools to data analysis.

Traditionally RDMS provides a powerful Structured Query Language tool to
query underlying DB. But it requires a full knowledge of DB schema as well
as explicitly specifying relation information in a query. Such approach is
used for creation of specific data-services or APIs which users can use,
but doesn't provide a way for end-users to place a query against
underlying DB. Recently we've been developed DBS-QL language \cite{DBSQL}
which demonstrates that queries can be placed against single DB in
flexible and intuitive way by end-users without any knowledge about
underlying schema. So it becomes desired to extend
this capability over other existing CMS DBs and have ability to
place queries across multiple data-services.

\section{Previous activity}
Cornell group has been played a significant role in design and implementation of
central CMS data bookkeeping system \cite{DBS}. 
%After original workshop
%hosted at Cornell in 2006 \cite{CornellWorkshop} we provided a roadmap 
%document \cite{roadmap} with guidelines to CMS DBS system. 
We also took a leading role in design and 
development of DBS query language \cite{DBSQL, DBS07} and CMS data discovery
service \cite{DD}.

\section{Data Aggregation Service}
Upon success of Data Discovery service \cite{DD} it was proposed
to extend this activity to other data services. A new project called
Data Aggregation Service was born, see \cite{DAS}. Under its unbrella
all meta-data can be queried using DAS Query Language similar to
DBS one \cite{DBSQL}. While such behavior is desired from  a user point of
view, its implementation represents a ``hot'' research topic in
Computer Science community, see \cite{Arms, DBXplorer, QueryAnswer, FedDB}.
Even though success of Information Retrieval (IR) and RDMS systems
are obvious, there is no a common way to utilize both sources of
information via simple, intuitive keyword search. Such demand has
been presented in various enterprise industries,
including health care, medical, financial domains.
While we cannot fully adopt keyword search (due to specific
domain features, such as find data who has bfield 0.4 and runs with
a specific data quality information), the adoption of DBS-QL
seems very reasonable choice.

%A keyword search over distributed relational and non-relational 
%data sources represents a demand in various enterprise industries, 
%including health care, medical, financial domains.

Upon preliminary studies, see \cite{DAS}, we explored a
system based on combination of RDMS and object-oriented databases
which will be capabable to provide desired requirements. Moreover,
it can serve as a cache buffer for variety of CMS data-services
and used by both production workflow tools and end-users.

We propose to design and develop DAS system based on existing
technologies, such as RDMS, IBM federated DB \cite{FedDB} or 
document-oriented databases \cite{CouchDB, MongoDB}, within
domain of CMS experiment and existing workflow APIs. Initially
the following data-services will be cosidered to be participants in
DAS system: data bookkeeping system, data location system, run summary, 
dashboard, data quality and luminosity DBs.
Each data-service will be gradualy added to DAS extending DAS
QL and amount of data processed and cached by DAS. The system will
undergoes to the analysis and testing, via DAS analytics, to address pre-fetching
strategies and scalability issues.

\section{Estimates}
A design and guidelines of DAS system has been summarized on
\cite{DAS} twiki page. Here we represent a main activities within
DAS project.

\begin{itemize}
\item
{\bf DAS core}: 1 FTE\\
Design and implementation of DAS core system. It comprises of the following
components:
\begin{itemize}
\item implementation of DAS request server
\item implementation of DAS Mapping DB
\item implementation of cache plugins, including hot and raw-caches, access
to underlying caching layer, etc.
\item implementation of data-service plugins, e.g. DBS plugin, phedex plugin, etc.
\item implementation of data populator, pre-fetch strategies, etc.
\item administrative toolkit
\item stress and scalability testing
\item unit tests for DAS components
\end{itemize}

\item
{\bf Data-service components}: 0.5 FTE\\
Work on plugable modules to access data-services,
including DBS, Phedex, LumiDB, RunSummary, DQ services.
This work mostly requires coordination between CMS project
leaders, ensuring production of DAS compliant services and data formats or
developing appropriate wrappers to existing legacy data-services. 
It can be paralelized among data-service developers.

\item
{\bf DAS query language}: 0.2 FTE \\
Design and implementation of DAS QL (Query Language);
work on QL syntax and semantics, lexer, parser.
We need to evaluate existing technologies, such as ANTRL or
other lexer/parser tools. Design and implement DAS QL syntax and
semantics; provide mapping between existing data-service APIs and
DAS QL.

\item
{\bf Web interface}: 0.2-0.5 FTE \\
Design and develop web interface for DAS discovery service.
We need to follow CMS SLA guidelines \cite{SLA} on providing secure
and integrated web service within existing CMS infrastructure
on cmsweb site. We must use existing CMS webtools framework
and infrastructure.

\item
{\bf DAS analytics}: 0.2-0.5 FTE \\
Design and develop DAS analytics DB, based on analysis of
common CMS queries to different data-services.
This work includes studies of common CMS queries to
various data-services, DBS, Phedex, RunSummary, etc. As a result
a pre-fetch strategies to DAS should be provided.
\end{itemize}

\section{Hardware requirements}
Initial prototype has been proven to utilize a multi-core technologies
allowing to take a full power of existing hardware. In stress tests
the 8-core machine, 2.33 GHz/core and 16GB of RAM has been
used. We plan to use it as a DAS request/processing server. In addition
we need to allocate significant disk space for caching purposes.
For example, existing size of DBS system is 27 GB (tables) and 37 GB (index).
So we should rely on a few TB of cache for DAS, possible extandable over
time and load on a system.

\section*{References}
\begin{thebibliography}{9}
\bibitem{DBSQL}  A. Afaq, V. Kuznetsov, L. Lueking, D. Riley, V. Sekhri,
''The CMS Dataset Bookkeeping Service Query Language (DBSql)''
To be published in CHEP 2009 procedings.
%\bibitem{CornellWorkshop}
%https://wiki.lepp.cornell.edu/hep/bin/view/DBSWorkshop/

\bibitem{DBS} A. Afaq, et. al. ``The CMS Dataset Bookkeeping Service'', CHEP 2007 
%\bibitem{roadmap} https://twiki.cern.ch/twiki/pub/CMS/DBS-TDR/roadmap.pdf
\bibitem{DBS07} A. Dolgert, V. Kuznetsov, C. Jones, D. Riley, 
``A multi-dimensional view on information retrieval of CMS data'', CHEP 2007
\bibitem{DD} https://cmsweb.cern.ch/dbs\_discovery
\bibitem{DAS} https://twiki.cern.ch/twiki/bin/view/CMS/DMWMDataAggregationService
\bibitem{Arms}
C. R. Arms, W. Y. Arms, ``Mixed Content and Mixed Metadata 
Information Discovery in a Messy World'',
chapter from ``Metadata in Practice'', ALA Editions, 2004

\bibitem{DBXplorer}
Sanjay Agrawal, Surajit Chaudhuri, Gautam Das: DBXplorer: A System for
Keyword-Based Search over Relational Databases. ICDE 2002: 5-16

\bibitem{QueryAnswer}
Georgia Koutrika, Alkis Simitsis, Yannis E. Ioannidis: Pr\'{e}cis: The Essence of
a Query Answer. ICDE 2006: 69-78

\bibitem{FedDB}
L. Haas, E. Lin,
``IBM Federated Database Technology'', \\
http://www.ibm.com/developerworks/data/library/techarticle/0203haas/0203haas.html

\bibitem{CouchDB}
http://couchdb.apache.org/

\bibitem{MongoDB}
http://www.mongodb.org/

\bibitem{SLA} 
https://twiki.cern.ch/twiki/bin/view/CMS/DMWTServiceLevelAgreement

\end{thebibliography}

\end{document}