### 2010-05-04 doc/paper.tex
Work on Section 3, 4, 4.1
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1806 4525493e-7705-40b1-a816-d608a930855b

% Template for Elsevier CRC journal article
% version 1.0 dated 13 October 2009

% This file (c) 2009 Elsevier Ltd.  Modifications may be freely made,
% provided the edited file is saved under a different name

% This file contains modifications for Procedia Computer Science

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% This template uses the elsarticle.cls document class and the extension package ecrc.sty
%% For full documentation on usage of elsarticle.cls, consult the documentation "elsdoc.pdf"
%% Further resources available at http://www.elsevier.com/latex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The '1p' and 'times' class options of elsarticle are used for Elsevier CRC
\documentclass[1p,times]{elsarticle}

%% The `ecrc' package must be called to make the CRC functionality available
\usepackage{ecrc}

%% The ecrc package defines commands needed for running heads and logos.
%% For running heads, you can set the journal name, the volume, the starting page and the authors

%% set the volume if you know. Otherwise `00'
\volume{00}

%% set the starting page if not 1
\firstpage{1}

%% Give the name of the journal
\journalname{Procedia Computer Science}

%% Give the author list to appear in the running head
%% Example \runauth{C.V. Radhakrishnan et al.}
\runauth{}

%% The choice of journal logo is determined by the \jid and \jnltitlelogo commands.
%% A user-supplied logo with the name <\jid>logo.pdf will be inserted if present.
%% e.g. if \jid{yspmi} the system will look for a file yspmilogo.pdf
%% Otherwise the content of \jnltitlelogo will be set between horizontal lines as a default logo

%% Give the abbreviation of the Journal.
\jid{procs}

%% Give a short journal name for the dummy logo (if needed)
\jnltitlelogo{Procedia Computer Science}

%% Hereafter the template follows `elsarticle'.
%% For more details see the existing template files elsarticle-template-harv.tex and elsarticle-template-num.tex.

%% Elsevier CRC generally uses a numbered reference style
%% For this, the conventions of elsarticle-template-num.tex should be followed (included below)
%% If using BibTeX, use the style file elsarticle-num.bst

%% End of ecrc-specific commands
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{comma,round}

% \biboptions{}

% if you have landscape tables
\usepackage[figuresright]{rotating}

% put your own definitions here:
%   \newcommand{\cZ}{\cal{Z}}
%   \newtheorem{def}{Definition}[section]
%   ...

% add words to TeX's hyphenation exception list
%\hyphenation{author another created financial paper re-commend-ed Post-Script}

% declarations for front matter

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\dochead{}
%% Use \dochead if there is an article header, e.g. \dochead{Short communication}

\title{The CMS Data Aggregation System}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

%\author[vkuznet]{Valentin Kuznetsov\corref{cor1}}
\author[vkuznet]{Valentin Kuznetsov}
\address[vkuznet]{Cornell University, Ithaca, New York, USA}
\ead{vkuznet@gmail.com}

\author[evans]{Dave Evans}
\address[evans]{Fermilab, Batavia, Illinois, USA}
\ead{evansde@fnal.gov}

\author[metson]{Simon Metson}
\address[metson]{Bristol University, Bristol, UK}
\ead{s.metson@bristol.ac.uk}

%\cortext[cor1]{Corresponding author}

\begin{abstract}
Metadata plays a significant role in large modern enterprises, 
research experiments and digital libraries where it comes from many different 
sources and is distributed in a variety of digital formats. 
It is organized and managed by constantly evolving software using 
both relational and non-relational data sources. Even though we can apply
an information retrieval approach to non-relation data sources,
we can't do so for relational ones, where information is accessed via
a pre-established set of data-services.

Here we discuss a new data aggregation system which consumes, 
indexes and delivers information from different relational and 
non-relational data sources to answer cross data-service queries 
and explore metadata associated with petabytes of experimental data. 
We combine the simplicity of keyword-based search with the precision of RDMS
under the new system. The aggregated information is collected from various sources,
allowing end-users to place dynamic queries, get precise answers and 
trigger information retrieval on demand. Based on the use cases of the CMS experiment, 
we have performed a set of detailed, large scale tests the results of which 
we present in this paper.
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
Meta-data\sep data aggregation\sep information discovery\sep HEP.

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\end{keyword}

\end{frontmatter}

%%
%% Start line numbering here if you want
%%
% \linenumbers

%% main text
\section{Introduction}
The European Organization for Nuclear Research, known as CERN, plays a leading
role in fundamental studies of physics. It is also known as a place where
many innovations in the area of computer science were developed, e.g. World Wide Web.
Today, the Large Hadron Collider (LHC) at CERN is marking a new era of High Energy
Physics (HEP), promising to deliver a few PB of data each year. 
At this scale the ``information discovery'' within heterogeneous, distributed 
environment becomes a key ingredient of successful data analysis.
The data and associated meta-data are produced in variety of forms and digital formats.
They are stored and retrieved from relational and non-relational data-sources, such as 
RDMS systems, document oriented databases, blogs, twikies, file systems,
customized applications, etc. Working in such environment requires a lot 
of expertise and easy way to look-up desired information is always a necessity.
A well-known solutions, e.g. data-services, are tighten to a specific data 
sources and end-users are left with a manual task to bookkeep and 
relate information from them.

Here we present a work on Data Aggregation System (DAS) which provides
ability to search and aggregate information across different 
data-services. While it was designed for CMS High-Energy Physics 
experiment at LHC, it can be applied elsewhere.

The rest of this paper is organized as following. 
In section \ref{RelatedWork} we discuss related work in a domain of 
keyword search over relational data sources.
Section \ref{DataModel} describes the CMS experiment and its data model. In section
\ref{DAS} we present architecture of the DAS system, including discussion of its
various components. Finally, our results are summarized in section \ref{Results}.

%As was pointed out in \cite{Arms} a mixed content and 
%mixed meta-data and meta-data consistency should be considered as a whole in design 
%of the system to successful information discovery.

\section{Related Work\label{RelatedWork}}
Even though the idea of keyword queries over the relational database
%querying relation databases via keyword based search
%algorithms 
is not knew it is still under discussion in computer science domain.
A few alternative solutions has been proposed to address this issue
in last several years. In \cite{DBXplorer} the conjunctive keyword queries,
i.e. retrieval of documents that contain all query keywords, has
been discussed. It requires that all keywords are matched in a row from a
single table or joined tables. Authors of \cite{QueryAnswer} are made a step forward
and evaluated information around matched values. The results were presented as
customized templates to end-users. 
These and other related works, see \cite{DBXplorer, QueryAnswer} 
for more references, are based on some kind of the graph over the underlying 
database and generated a new database out of pristine one to perform the work.
The proximity of the results in those techniques are mostly confined to a 
text-based values stored in a database, since the usage of numerical ones 
is mostly useless without knowledge of the context of the input value. 
For example, typing a value 100 in the search input field can lead to 
plenty of inappropriate matches, e.g. row ids. In order to be
answered correctly it requires additional keyword, clarifying its meaning, 
which should be matched with particular entity (table.column) in the 
underlying database. Therefore the logical questions, e.g.
{\it find me total number of files whose size more then 20 but less then 
100} are hard to answer precisely using approaches outlined in \cite{DBXplorer, QueryAnswer}, 
while it can be easily accomplished using SQL.

To address those limitations a keyword-based query language was introduced in \cite{DBS-QL}.
It used a set of pre-defined keys mapped to underlying schema and shortest path
algorithm which employed a foreign key constrains to build SQL queries. As a result
the question
{\it I'm looking for files who contain data taken on certain date and located at
particular site} was represented as simple as \cite{DBS-QL}
\begin{verbatim}
find file where date 2009-01-02 11:59 CET and site = T2
\end{verbatim}
without specifying any relational details.
This solution represents a common use case in HEP community and provides
an intuitive mapping between mental model of the end-users and 
underlying database schema.

At the same time the data aggregation is commonly addressed in the context of
federated databases, see for example \cite{FedDB}. The data from different 
RDMS systems are stored into the federated DB where SQL queries can be used to 
search desired data. Apart from many internal obstacles, e.g.
schema, domain and naming conflicts, the external ones are often
a barrier to work with in distributed, heterogeneous environment where
different security policies are in place. At the end an usage of SQL and/or
any other Query Language will be required to search for data.
%Even though it would be interesting to apply technique 
%discussed in \cite{DBS-QL} to a federated DB it still requires 
%an access to the underlying database schema.

To avoid those limitations we decided to build a keyword-search based system around
existing data-services whose nature and access policies are intact.
%The information retrieval were triggered on demand and stored into the cache.
The key ingredients were the setup of the mapping and analytics services which provided
a necessary glue across existing API/notations and user queries. This approach 
allows us to perform dynamic data aggregation across various data-services, achieve
precise answers using keyword-based queries
and organize information retrieval on demand.
%to store meta-data information 
%We wanted to expand this approach further and apply it 
%to handle cross data-service queries from broad variety 
%of relational and non-relational data-sources within 
%our distributed heterogeneous environment.

\section{CMS data model\label{DataModel}}
The CMS, Compact Muon Solenoid, \cite{CMS} 
is one of the two particle physics detectors built on 
the proton-proton Large Hadron Collider (LHC) at CERN in Switzerland and France. 
It is designed to explore the frontier of the HEP and provide physicists
the ability to look at the conditions presented at the early stage of our Universe.
More then 3000 physicists from 183 institutions representing 38 countries 
are involved in the design, construction and maintenance of the experiment.

The CMS distributed computing and data model \cite{CMSDataModel} 
is designed to process
and efficiently manage a few PBs of data expected each year
in CMS operation at LHC. The computing resources provided by members of CMS
collaboration are geographically distributed, 
interconnected via high throughput networks and operated by means 
of the Grid software. The model allows to cover broad variety of
hardware, mass and storage elements and configuration of the
clusters. To accommodate efficient data processing CMS uses
a multi-Tier distributed model, where specific tasks of data taking,
e.g. processing, archival and distribution, are assigned to each tier based
on CMS data model. For example, the Tier-0 center at CERN is responsible
for archiving the data coming out from the detector, prompt first pass reconstruction
and data distribution to Tier-1 centers, located around the world.
%The 7 Tier-1 centers
%located in France, Germany, Italy, Spain, Taiwan, United Kingdom and United States
%keep a portion (copy) of the data delivered by Tier-0 for further processing.
They provide storage and CPU power for high priority analysis jobs.
The Tier-2 centers
%, located at more then 50 sites around the world,
are dedicated for user analysis tasks and production of simulated data.

A broad variety of data-services have being designed and developed to
maintain detector and production operations, including detector
conditions databases, data location and bookkeeping services,
data transfer and job monitoring tasks. Even though majority of them
are located at CERN, it was never been a requirement in CMS computing
and data model. 
%For instance, the production teams operated at Tier-1,2
%centers set up local services for data bookkeeping and operational
%tasks. Based on their resources the choice of database back-end was
%up to the site managers. Therefore the CMS software were designed to support different
%DB back-ends and provide tools for migrating data across them.

Once such conglomerate of data-services starts operating the obvious
question arises: how to find out desired information across multiple data-services
in our distributed environment? Even though individual data-services were designed
to answer specific questions about the data they serve, the ability to search and relate
information among them was always a tedious human task. The growing amount of information
and desire to make cross-service queries force us to design and develop new
Data Aggregation System.

\section{Data Aggregation System\label{DAS}}
The goal of the Data Aggregation System was to fetch and aggregate meta-data 
on demand from existing CMS data-services under one umbrella by using their APIs, 
security policies, etc.
In contrast, the approaches
discussed in \cite{DBXplorer, QueryAnswer, FedDB} are
enforced a publication step of data into a dedicated database 
or central repositories \cite{iRODS}. Therefore, quite often it was required
that data should be converted into common data structure/format in order to enable
aggregation, see \cite{OpenArchive}. In DAS the data are provided
by existing data-services and infrastructure, whose implementation,
data representation and security policies are left intact. Consequently,
DAS does not require a data preservation and transaction capabilities. 
Instead DAS can be treated as proxy or caching layer where information 
retrieval and aggregation are triggered dynamically by user queries. 
Below we discuss in details the DAS architecture.

\subsection{DAS architecture}
The design of the DAS system was based on previous studies of CMS Data 
Bookkeeping System (DBS) \cite{DBS, DBS07}. The DBS Data Discovery service
\cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provided
a syntax similar to SQL, without specific knowledge of underlying DB schema.
%We carefully analyzed user
%requests and queries placed into the DBS, their patterns, frequencies and latency. 
%As a result we created a data discovery service \cite{DD}. Originally it used MVC
%architecture and later, being re-factored as pure presentation layer on top of 
%the DBS Query Language (DBS-QL) \cite{DBS-QL} with support of dynamic queries against
%DBS back-ends. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language, where 
%we took away relations and details on underlying schema from end-users. 
%This was achieved by using Dijkstra's shortest path algorithm 
%to establish necessary table joins based on provided information in user queries and
%knowledge of foreign key relationships from auto-loaded database schema.
%The DBS-QL keys exposed to end-users were chosen based on a jargon used by
%physicists. They're internally mapped into database entities and attributes, 
%e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
A quick adoption and wide usage of DBS-QL in CMS gave us confidence in chosen approach 
and provide a basis for building DAS system.

As was pointed out in \cite{Arms} a mixed content and 
mixed meta-data and meta-data consistency should be considered as a whole in design 
of the system to successful information discovery. 
Starting from this ground we are designed DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support a keyword based search queries with ability to use condition operators;
%\item pluggable interface to existing and up-coming data-service APIs;
\item use existing heterogeneous software environment and distributed nature of the data-services;
\item preserve access rules and security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementation and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. 
Input query were placed into analytics DB and mapped into a set of
data-service APIs and their parameters. If superset of the query has been
found in DAS cache, the appropriate results were delivered back to the users.
Otherwise a data-service APIs were invoked, triggering DAS cache population. 
Once data become available and aggregated in DAS cache, users were able 
to see the results. This on demand system was supplemented by DAS robots 
(Unix daemons) who were consulting DAS analytics database to enable data 
population into DAS cache for most common query requests.
}
\label{DAS_cache}
\end{figure}

\noindent
The choice of existing RDMS as DAS back-end(s) was ruled out for several reasons. 
For instance, we didn't require any transactions and data persistency in DAS;
a dynamically typing of stored meta-data objects was one of the requirements, etc.
Those arguments forces us to look for alternative IT solutions.
Through the careful analysis of available options, such as a file-based and in memory caches, 
key-value databases, documented-oriented databases we made our choice in favor 
of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
MongoDB \cite{MongoDB}. Both systems are ``schemaless'' databases providing
arbitrary document structure storage, replication and fail-over features. 
We made our decision in favor of MongoDB, due to its support of dynamic queries, 
full indexes, including inner objects and embedded arrays,
and auto-sharding. Our preliminary benchmarks shown that it can sustain
the desired load and size of our meta-data information. We used 
document-oriented database, MongoDB, as a back-end for three 
main DAS components: Mapping, Analytics and DAS cache databases, 
which will be discussed next.

The DAS architecture is shown in Fig. \ref{DAS_cache}. It consists of
the core library with support of pluggable modules for data retrieval;
caching layer to store and aggregate results;
request service to handle user queries;
Mapping DB to keep information about data-service APIs, e.g.
notations and mapping to DAS keys;
Analytics DB for query analysis.

\subsection{DAS core}
The DAS core communicated with different data-services and retrieved
data on demand. This was done by invoking appropriate data-service plug-in(s)
upon the user request. Each plug-in consists of DAS map, the
data parser(s) to convert data-service output into DAS records
and it was responsible for data-service API calls.
The data-service plug-in contains a map about data-service APIs,
input and output parameters and their mapping to DAS keys. For example
\begin{verbatim}
{"api":{"params":{"name":""}, "name":"CMSNametoSE"}, "system":"sitedb",
 "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
              "api_param": "name", "das_key": "site"}], 
 "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
              "key": "site", "map": "site.name"}]}
\end{verbatim}
This information was stored into DAS Mapping DB which we discuss next.

Although the majority of data-service plug-ins looks alike
and their logic was abstracted, including parsers, there were a few
cases where special care was required, e.g. access to data-services 
accessible via private network. Therefore, a way to propagate 
user credentials was required, leading to specific plug-in implementation.

%All user queries were written to Analytics DB for further analysis, while
%mapping between data services were kept in Mapping DB. The data-service 
%output were re-mapped into DAS notations and, if necessary, aggregated 
%inside of the DAS cache.

\subsection{DAS Mapping DB}
The Mapping DB hold the mapping between data-service APIs and DAS keys exposed to
the users. It created a necessary glue between data-services and identifies
how their data can be related to each other. Basically Mapping DB was positioned 
as a {\it registration} service with ability to translate notations among data-services.
Upon identification of APIs which will participate in DAS activity 
we stored their names, input parameters, together with their type and accept 
values, as regular expression patterns, into the Mapping DB. 
Also we stored the mapping between data-service notations and DAS keys. 
For instance, the CMS DBS system \cite{DBS}
used a \verb+logical_file_name+ notation to identify the name of the
logical file, while the same entity was called \verb+lfn+ 
in CMS PhEDEx system \cite{PhEDEx}.
Since both of them were referred to the logical file name, 
we map both notations into DAS key \verb+file+.
%\footnote{We
%also deal with physical file names, but still used {\it file} to refer to them in
%DAS notations, since simple regular expressions were used to identify the
%meaning of the {\it file} keyword from the provided value.}
Such keys, e.g. \verb+file+, were exposed to the end-users as pre-defined keywords to
be used in search queries. For instance, end-users were able to type \verb+file+ in
order to get all file names known to DAS via data-services as well as
provide a concrete patterns associated with a file, e.g. \verb+file = abc*+.
Under the hood, the appropriate data-service APIs were looked up along with appropriate
parameters to answer those questions. In aforementioned example, 
usage of \verb+file+ in DAS query cause a call of DBS 
\verb+listFiles+ and PhEDEx \verb+fileReplicas+ APIs, respectively.
This approach gave ability to end-users to express the meaning of their queries, especially
when they deal with numbers, and achieve accuracy of results. 
For instance, users were able to type \verb+run > 10 and run < 100+
to specify a range of run numbers to be looked up in DAS cache. At the same time
a full keyword search is under development.
% within MongoDB.

\subsection{DAS Analytics DB}
The DAS Analytics DB played a special role. It collected information
about user requests placed to the system. Each request was recorded. Upon its
decomposition into set of selection keys and conditions we also recorded
API name and input parameters into the Analytics database and 
keep updating their counter information for repeated calls. 
This allows to keep track of frequency of API calls and establish 
pre-fetch strategies for most common queries.

\subsection{DAS caching system}
The DAS caching system was used to dynamically fetch
and aggregate data upon user requests. We used the following 
algorithm:
\begin{enumerate}[1.]
\item Decompose query into set of conditions (\verb+conds+) 
and selection keys (\verb+skeys+)
\item Check cache for query hash, if found, look-up results from the cache
\item Identify list of APIs which can serve requested data
\begin{verbatim}
apilist = [r(url, api, args) for r in mapping_db(skeys, conds)]
\end{verbatim}
Here \verb+mappind_db+ defines translation between input query and data-service
APIs which can delivery such data.
\item Iterate over APIs
\begin{verbatim}
for url, api, args in apilist:
    if  superset_in_analytics(api, args):
        continue
    for row in parser(get_data(url, api, args)):
        yield row
\end{verbatim}
Here \verb+superset_in_analytics+ verifies if provided \verb+api, args+
are subset of already registered in Analytics DB api call(s). The 
\verb+get_data+ and \verb+parser+ are define methods to fetch and parse
data, respectively.
\item Perform aggregation by iterating over records from the previous step
\begin{verbatim}
if cache_has(row.primary_key):
   existing_record = get_from_cache(row.primary_key)
   merge(row, existing_record)
   insert_into_cache(row)
   delete_from_cache(existing_record)
else:
   insert_into_cache(row)
\end{verbatim}
The \verb+primary_key+
is a key which identify the record defined in Mapping DB, e.g. \verb+file.name+
for \verb+file+ record.
We did use a bulk insert operations at this step to increase throughput. 
\item Get results from the cache
\end{enumerate}
%The data was stored in a form of DAS records into the cache, while
%queries were recorded into Analytics DB for further analysis.
We also had an option to run independent set of daemons, DAS robots, who
monitor Analytics DB for most popular requests and pre-fetch the most
common data into DAS cache. 
%The interesting part was being able
%to identify the superset of api/args out of found one in an Analytics DB.

%The DAS caching system was used to store data-service results and aggregated
%information in form of DAS records. As it shown in figure \ref{DAS_cache} 
%the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
%and aggregate data over there. The independent set of daemons, DAS robots, were
%constantly monitor Analytics DB for most popular requests and pre-fetch most
%common data into DAS cache.

%Each user request placed to the system was parsed and set of selection
%and condition keys were identified. They were mapped into set of
%data service APIs and their input parameters. DAS
%cache was checked for a queries with a similar set of selection and condition
%keys used earlier.\footnote{Since DAS didn't hold
%any data by default we cannot rely on data being present in a cache.
%Moreover if we look-up data first, we cannot tell if it was complete
%set or not, since earlier queries may retrieve only portion of the requested data.}
%If the superset was found a data were
%retrieved from the cache, otherwise a data-service APIs were
%invoked and the results were collected and aggregated inside of DAS cache.

\subsection{DAS queries}
Apart from the previous DBS-QL syntax we decided to drop its syntax and allow users
express their queries in a free text-based form. Although we must clarify its
meaning. We used a pre-defined set of keys registered in DAS to describe common
entities of our data. For example, \verb+dataset+, 
\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
was based on analysis of daily work operations of our users, see \cite{DBS07}.
Each key had a set of attributes with primary one assigned to the key itself.
For instance, the \verb+block+ key had the following attributes
\verb+name+, \verb+size+, \verb+replica+ etc., 
with primary attribute being \verb+name+.
The choice of attributes were based on a set of meta-data information
associated with a given DAS key. Therefore end-users where able to specify
a context in their queries by using appropriate attribute(s). For example,
for query \verb+block+ we didn't perform a string matching to the
``block'' string, but rather look-up a block objects in a DAS, while query
\verb+block.name=abc*+ instructs system to look-up appropriate blocks
with matched names. At the same time, to complement the strength of the 
DAS, we do plan to index dynamically retrieved meta-data and add a keyword-based 
search later.

The free text-based queries were transformed into MongoDB query syntax, 
who was used internally in DAS. It represents a set of selection 
keys and conditions as a JSON document. For example, a query
{\it dataset site=T1\_CH\_CERN}
was transformed into 
{\it \{"fields" : ["dataset"], "spec" : \{"site" : "T1\_CH\_CERN"\}\}}.
Such syntax was perfect fit for us, since it allows to store all queries
as JSON documents into Analytics DB for further analysis. It is worthwhile to note that
MongoDB query syntax is very reach. It supports advanced queries with 
boolean expressions, conditional operators, regular expressions, etc. providing
a real power for our end-users.

\subsection{Data-services and aggregation}
As we outlined above the DAS Mapping DB becomes an authorative
source of information about all data-services participating in DAS.
Based on information stored in Mapping DB we were able to 
retrieve appropriate data from different data-services, re-map them into 
DAS notations and aggregate them on demand. For instance, the Data 
Bookkeeping System (DBS) and data location system (PhEDEx)
provided information about CMS block meta-data. In former case, DBS 
stored information about blocks and their relationships to other 
physics objects, such as dataset and files, while PhEDEx kept
information about its location and replicas. Upon a user request to 
find a block meta-data, we were able to identify appropriate set of DBS and PhEDEx APIs, 
retrieve data from them, re-format their output to DAS notations 
and store results into DAS cache performing aggregation of matched 
objects using \verb+block.name+ key. The merged results were stored 
as a set of data records. Here is an example of such aggregated record:
\begin{verbatim}
{"das_id": ["4b195404e2194e21c4000030", "4b1953fee2194e21c4000002"], 
 "block": [{"name": "/a/b/c#123", "replica": [{"complete": "y", 
            "subscribed": "n", "custodial": "n", "site": "T1_US_Buffer", 
            "nfiles": "18", "se": "srm.site.com", 
            "size": "112694171838"}, {....}]},
           {"name": "a/b/c#123", "nevents": "570200", 
            "created_by": "USER_DN", "path": "/a/b/c", ...}],
 "das": [{"expire": 1259952908}, {"expire": 1259954702}]}
\end{verbatim}
Here \verb+das_id+ refers to api call records stored in Analytics DB,
\verb+block+ represents an aggregation block meta-data information from DBS
and PhEDEx systems and \verb+das+ contains their expiration time stamps.
Usage of this information was valuable contribution in debugging process of
data-services themselves, e.g. identification of not-matched records, 
latency studies, etc.
% Here is an example of DAS record representing
%aggregation information between two SiteDB APIs, {\it CMStoSAMName} 
%and {\it CMStoSiteName}, with API results merged under the {\it site} key.
%\begin{verbatim}
%{"site":[{"samname": "CERN-PROD", "name": "T1_CH_CERN"}, 
%         {"sitename": "CERN", "name": "T1_CH_CERN"}], 
% "das": [{"ctime": 0.9, "expire": 1256362175, "api": ["CMStoSAMName"],
%          "url": "https://a.b.com/sitedb/json/index/CMStoSAMName", 
%          "timestamp": 1256318975, "system": ["sitedb"]},
%         {"ctime": 1.2, "api": ["CMStoSiteName"], "expire": 1256362174},
%          "url": "https://a.b.com/sitedb/json/index/CMStoSiteName", 
%          "timestamp": 1256318974, "system": ["sitedb"]]}
%\end{verbatim}

%We used RESTful model \cite{REST} for all DAS cache server APIs.
%It allows us to share core library and other components during development of 
%various DAS clients, e.g. CLI, web interface, robots, etc. and hides
%complexity of access to individual data-services and their security policies
%as well as proper organize caching solution.
%For instance, to access on-line Run Summary information we were forced to use
%Grid certificates. Instead of propagating user certificate to data-service
%the DAS cache server works as a proxy to this server and used host certificate 
%to serve such requests on behalf of the users.

\section{Results\label{Results}}
In this section we discuss preliminary results achieved with DAS system
using the following CMS data-services:
\begin{itemize}
\item Data Bookkeeping System, DBS \cite{DBS}, the CMS meta-data service;
\item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
CMS file transfer system;
\item SiteDB \cite{SiteDB}, the CMS site resources data-service;
\item RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
%\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
%\item DQ \cite{DQ}, Data Quality data-service which collects information
%about detector conditions during data taking;
%\item Overview \cite{Overview}, collects information about CMS
%transfer rate, etc.;
\item Dashboard \cite{Dashboard}, the LHC virtual organization
monitoring data-service portal;
\end{itemize}
Each data-service has its own scope, size and peculiar properties. 
For instance, the data in PhEDEx are transient due to constant 
migration of CMS data; the DBS system were divided into dozen 
of individual instances with own size and lifetime, who has been used by different production 
teams for various bookkeeping tasks; the RunSummary data-service
operates on private network accessible at CMS online operation center.
%Table \ref{db_size} shows the current size of meta-data from aforementioned 
%data-services.
%\begin{figure}[htb]
%\centering
%\includegraphics[width=100mm]{db_size.pdf}
%\caption{
%Current size of meta-data for CMS data-services participated in DAS, before
%official data-taking. 
%}
%\label{db_size}
%\end{figure}

%\begin{table*}[hbt]
%\centering
%\begin{tabular}{llllll}\hline
%\hline

%Size of  & Dashboard & DBS & PhEDEx & SiteDB & RunSummary \\
%\hline
%tables (GB) & 40.8  & 28.0 & 5.5 & 0.006 & 0.9 \\
%indexes (GB) & 135.9 & 57.2 & 12.6 & 0.004 & 2.7 \\
%\hline
%\hline
%\end{tabular}
%\caption{
%Current size of meta-data for CMS data-services participated in DAS, before
%official data-taking. 
%}
%\label{db_size}
%\end{table*}
At the moment, the amount of meta-data stored in all services is equal to
$\sim$75 GB and $\sim$200 GB in tables and index sizes, respectively. While 
we estimate to accumulate $\sim$500 GB of the meta-data each year during 
normal CMS operation at LHC.
%\footnote{This number is based on tests 
%performed in CMS during Cosmic Run at Full Tesla (CRAFT) data taking \cite{CRAFT09}.}
%In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
%billion events, 330TB in 32 datasets:
%AlCa/Calibration RAW: 926 million events, 29TB
%Physics RAW: 1.24 billion events, 147TB
%PromptReco: 1.24 billion events, 111TB
%Bulk ALCARECO: 2.34 billion events, 11TB
%Express FEVT: 105 million events, 24TB
%Express ALCARECO: 153 million events, 0.99TB
%HLTMON FEVTHLTALL: 31 million events, 7.4TB
%Above event numbers have significant overlap by design, 
%2.2 billion unique events, including 524 million cosmic muon
%triggers
% 0.3PB - 50GB meta-data in DBS
% 3PB/y - x GB

To test DAS performance we measured elapsed time and CPU/RAM utilization
required to collect and aggregate information from different
data-services. Table \ref{DAS_benchmark} summarizes the biggest 
contributors so far, the block meta-data information provided by DBS and 
PhEDEx CMS data-services.

\begin{table*}[hbt]
\centering
\begin{tabular}{lllll}\hline
\hline
System & Format & Records & Elapsed time & Elapsed time \\
& & & no cached data & w/ cached data \\
\hline
DAS (DBS) & XML & 382396 & 63 sec & 0.98 sec \\
DAS (PhEDEx) & XML & 186145/180439 & 164 sec & 0.98 sec \\
DAS (aggregate) & JSON & 388108 & 234 sec & 2.05/50.7 sec \\
\hline
\hline
\end{tabular}
\caption{Time required to fetch, parse and aggregate block information
from DBS and PhEDEx systems. In the case of PhEDEx
system we show total number of fetched records 186145 together with
number of records merged in DAS 180439. The total number of DAS records 
were calculated as number of aggregated records plus left overs,
the non-matched records from both systems. The look-up time is presented in
last column, where for DAS it shows as query look-up time (2.05 sec) 
and time to get all records to disk (50.7 sec).}
\label{DAS_benchmark}
\end{table*}

The initial DAS implementation was done in Python language \cite{Python}
and all tests were performed on x64 Linux platform where
1 CPU was used by DAS server and 1 CPU by MongoDB database server. 
%The CPU utilization
%of DAS server was 67\% in average and RAM usage was below 35MB. 
All tests were performed multiple times to avoid fluctuations. 
The elapsed time measured during those tests consists of the 
following components:
\begin{itemize}
\item[]
{\it
Elapsed time = retrieval time + parsing time + re-mapping time 
        + cache insertion/indexing time 
        + (aggregation time) + (output creation time)
}
\end{itemize}
Here the {\it retrieval time} is a time required to access data from remote data-service,
{\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
is a time required to convert data-service meta-data to DAS records for every object
we parse, {\it cache insertion} and {\it indexing time} represents time spent on DAS 
cache back-end, {\it aggregation time} is required to merge objects into DAS records based
on their common key (\verb+block.name+ in this case) and {\it output creation time}
is a time required to write DAS records to disk. Please note that last two
components, {\it aggregation} and {\it output creation time}, were only applied to
one of the sub-system, either DBS or PhEDEx, depending on insertion order.
% since the first one was injected into
%the fresh database, while the second required the merging step.

%DAS execution time (phedex) 160.606897116 sec
%DAS execution time (dbs) 52.5349619389 sec
%DAS execution time 256.0670228 sec, Sat, 28 Nov 2009 15:46:39 GMT
%
%DAS execution time (phedex) 0.952513933182 sec
%DAS execution time (dbs) 0.951184034348 sec
%DAS execution time 44.7982931137 sec, Sat, 28 Nov 2009 15:52:11 GMT

The individual tests of DAS components (DBS and PhEDEx) shown roughly 
the same performance. Although the time of PhEDEx component was bigger 
due to longer retrieval and parsing time.
The time of PhEDEx component shown in Table \ref{DAS_benchmark} is divided 
among retrieval and parsing, $\sim$110 sec, and merging steps, $\sim$50 sec,
respectively. As shown in Table \ref{DAS_benchmark},
the look-up time spent to access individual components (last column) was quite
reasonable, roughly 1 second for both systems. While final time to
get DAS records on disk was about 50 seconds. This was mostly due to I/O and
conversion operations between binary data format used by MongoDB, BSON, to DAS
data format, JSON.

Based on results of stress tests we measured that we can achieve throughput of
$\sim$6000 docs/sec for insertion, $\sim$1100 docs/sec for aggregation and 
$\sim$7600 docs/sec for 
reading operations in DAS, which are suitable for our needs. Since all 
operations were performed on a single CPU a further improvements are expected. 
Also, a necessary tuning can be done by using C++ driver instead of python one.

\section{Summary}
We presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
at LHC, CERN, Geneva, Switzerland. It was designed to provide caching and
aggregation layer on top of the existing relational and non-relation data-services
mostly in real time fashion. All of the data were retrieved on demand basis,
while data pre-fetching of most common queries is in our plans. We developed
prototype in python for DAS system which is under commissioning phase right now 
and presented performance studies in this paper. 
The CMS experiment is started to collect its data in December of 2009 and 
we expect to accumulate a few PB of real data each year. 
In addition, the Monte-Carlo samples will be produced at this scale.
Based on test studies performed in CMS, we expect that total size of
meta-data produced each year by experiment will be of the order of
500GB. The DAS should sustain such load and will provide generic 
``data-discovery'' service for CMS experiment in years to come.

\section{Acknowledgments}

This work was supported by the National Science Foundation and 
Department of Energy of the United States of America. 
Fermilab is operated by Fermi Research Alliance, LLC under Contract
No. DE-AC02-07CH11359 with the United States Department of Energy.

%% References
%%
%% Following citation commands can be used in the body text:
%% Usage of \cite is as follows:
%%   \cite{key}         ==>>  [#]
%%   \cite[chap. 2]{key} ==>> [#, chap. 2]
%%

%% References with BibTeX database:

%\bibliographystyle{elsarticle-num}
%\bibliography{<your-bib-database>}

%% Authors are advised to use a BibTeX database file for their reference list.
%% The provided style file elsarticle-num.bst formats references in the required Procedia style

%% For references without a BibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have the following form:
%%   \bibitem{key}...
%%

% \bibitem{}

% \end{thebibliography}

\section*{References}
\begin{thebibliography}{00}
\bibitem{DBXplorer}
S. Agrawal, S. Chaudhuri, G. Das,
``DBXplorer: A System for Keyword-Based Search over Relational Databases'',
ICDE 2002, pp. 5-16.

\bibitem{QueryAnswer}
G. Koutrika, A. Simitsis, Y. E. Ioannidis,
``Pr\'{e}cis: The Essence of a Query Answer'',
ICDE 2006, pp. 69-78.

\bibitem{DBS-QL} 
V. Kuznetsov, D. Riley, A. Afaq, V. Sekhri, Y. Guo, L. Lueking,
``The CMS DBS Query Language'', CHEP 2009.

%\bibitem{AMI}
%Altas AMI web portal,
%http://ami.in2p3.fr/opencms/opencms/AMI/www/Tutorial/AMIMediation.html

\bibitem{FedDB}
L. Haas, E. Lin,
``IBM Federated Database Technology'', \\
http://www.ibm.com/developerworks/data/library/techarticle/0203haas/0203haas.html

\bibitem{CMS} 
R. Adolphi et al., 
``The CMS experiment at the CERN LHC'',
JINST 0803, S08004 (2008).

\bibitem{CMSDataModel} 
C. Grandi, D.Stickland, L.Taylor et al.,
``The CMS Computing Model'',
CERN-LHCC-2004-035/G-083 (2004).
%A. Fanfani et. al., 
%``Distributed Analysis in CMS'', 
%to be published in Journal of Grid Computing.

\bibitem{iRODS}
The Integrated Rule-Oriented Data System,
http://www.irods.org/

\bibitem{OpenArchive}
Open Archives Initiative,
http://www.openarchives.org/

\bibitem{DBS} 
A. Afaq, et. al.,
``The CMS Dataset Bookkeeping Service'', 
J. Phys. Conf. Ser, 119, 072001 (2008).

\bibitem{DBS07} 
A. Dolgert, V. Kuznetsov, C. Jones, D. Riley, 
``A multi-dimensional view on information retrieval of CMS data'',
J. Phys. Conf. Ser, 119, 072013 (2008).

\bibitem{DD} https://cmsweb.cern.ch/dbs\_discovery

\bibitem{Arms}
C. R. Arms, W. Y. Arms,
 ``Mixed Content and Mixed Metadata 
Information Discovery in a Messy World'',
chapter from ``Metadata in Practice'', ALA Editions, 2004.

\bibitem{MySQL}
http://www.mysql.com/

\bibitem{CouchDB}
http://couchdb.apache.org/

\bibitem{MongoDB}
http://www.mongodb.org/

%\bibitem{REST}
%Fielding, Roy Thomas ``Architectural Styles and the Design of 
%Network-based Software Architectures'', Doctoral dissertation, 2000,
%University of California, Irvine

\bibitem{PhEDEx}
%Rehn et. al.,
%``PhEDEx high-throughput data transfer management system'', CHEP 2006;
%R. Egeland et al., 
%``Data transfer infrastructure for CMS data taking'', 
%Proceedings of Science, PoS (ACAT08) 033 (2008);
L. Tuura et al., 
``Scaling CMS data transfer system for LHC start-up'', 
J. Phys. Conf. Ser, 119, 072030 (2008).

\bibitem{RunSummary}
https://cmswbm.web.cern.ch/cmswbm/cmsdb/servlet/RunSummary

\bibitem{SiteDB}
https://cmsweb.cern.ch/sitedb/

%\bibitem{LumiDB}
%https://twiki.cern.ch/twiki/bin/view/CMS/CMS-DMWM-DBS-Luminosity

%\bibitem{DQ}
%Data-Quality DB reference

%\bibitem{Overview}
%https://cmsweb.cern.ch/overview/

\bibitem{Dashboard}
J. Andreeva, et. al,
``Experiment Dashboard – The Monitoring System for the LHC Experiments'',
GMW’07, June 25, 2007, Monterey, California, USA.

%\bibitem{CRAFT09}
%https://twiki.cern.ch/twiki/bin/view/CMS/CRAFT09AnalysisInfo

\bibitem{Python} Python programming language, http://www.python.org

%\bibitem{DBSearch}
%D. Konopnicki, O. Shmueli,
%``Database-Inspired Search'', 
%Proc. of the 31st VLDB Conference, 2005.
\end{thebibliography}
\end{document}