### 2012-10-19 doc/thesis/latex/proposed_solutions.tex

\section{Proposed solutions}
\subsection{Ease of use}

\begin{itemize}
\item Interactive guidance on how to compose queries to ease the learning (could also include some examples of popular queries):
\begin{enumerate}
\item What entity are you searching for?
\item How could you identify it (i.e. what do you know)?
\item What do you want to do with these results (after seeing the results)? Possible options: list items; select only specific columns; do aggregation; use in command line/another program
\end{enumerate}

\item JavaScript-based query interpreter with auto completion to ease query writing: could suggest searchable entities and maybe even entity attribute names.

\item Given a keyword query, suggest the best matching structured queries (as a ranked list)%
 	\footnote{As DAS uses Boolean IR model we could only rank specific structured queries, but not the results itself.}:
	\begin{itemize}
	\item the algorithm will generate mappings between each keyword to either an entity name, selection attribute name or its value (resembling 'Query forms' and 'Keyword cleaning' approaches)
	\item ranking could be based on: how closely keywords are matching (some or all) required API's input parameters, popularity of certain queries, users' feedback	 and user's personal profile (e.g. search history, department, etc)
	\item build an inverted index of attribute values with some good full-text search toolkit%
				\footnote{e.g. Xapian (http://xapian.org/) that is known for its flexibility, or Sphinx (http://sphinxsearch.com/)}. 
	If no match found (e.g. new entry not yet in our cache) pattern matching could be used to make a best guess
	\item the system could improve its results from users' feedback through machine learning (e.g. HMM)
	\item multi-wildcard queries (e.g. dataset=*Zmm*special*RECO*) are common at CMS that will need special attention\footnote{%
		Matching a wildcard query into a schema field will be harder. Also for performance reasons some web-services will introduce limitations on the format of wilcard queries, while DAS wants to allow more flexible queries, at least for \textit{datasets}}.

\item knowing all possible APIs' parameter values, in addition to historical values seen so far, would improve quality of the keyword search (better mapping from keywords to these values) or even the structured search e.g. through query cleaning \& auto-completion. {\color{red}Not yet known if available}%
		\footnote{problem: the owners of services may not want to provide direct access to DB because of no trust, security or performance issues. Otherwise, specific APIs could be developed to return possible parameter values but this is quite an overhead}.
\begin{itemize}
\item having direct access to DB, one could incrementally index target database\footnote{%
			E.g. an inverted index of values in Oracle/MySQL DB tables can be built to keep the list of possible terms. For instance, Sphinx full text search engine can access DBs directly and also supports incremental indexing but it needs a bit of manual configuration {http://sphinxsearch.com/docs/current.html\#delta-updates} }
	tables semi-automatically mapping it's columns into API parameters.

%\item may need to exclude very large tables
\end{itemize}
	\end{itemize}
\end{itemize}

\subsection{Supporting complex queries}
{\color{red}At the moment this still needs some further analysis/consideration, however it seems even the existing data-source definitions could be used for building more complex queries.}

DAS Service definitions contain: for each data service (API), mapping of its parameters and result type into entities of mediated schema (but NOT ALL their fields are defined a priori). Even if DAS definitions do not describe every attribute of API result, as all input parameters are attributed to a mediated schema entity, these definitions seem mappable  into \textit{Information Manifold's} data source definitions described on Sec.\ref{IM_query_translation} of Literature review.

Other issues to consider:
\begin{itemize}
\item how these queries shall be entered by the user and how ambiguity shall be handled especially if queries entered in some simplified (semi-structured) language like DAS Query Language currently is
\item any optimizations/heuristics if multiple possible solutions exist
\item handling complex and heavy queries with large number of results, e.g. lazy loading with pagination (at least at the level of service composition as APIs do not yet support pagination/sorting)
\end{itemize}

\subsection{Performance}

%%%%%  -------------------------------------------

\subsubsection*{More intelligent caching}
Some of the data instances changes very rarely. For example in DBS system old datasets would never change, while new entities are constantly added (some of their attributes may change, but only rarely, e.g.  validity of a dataset). 
%
Cached copy may be shown by default while up to date results could be retrieved on user's request. An automatic change rate prediction could be useful to efficiently balance between caching and retrieving results.
%
On the other hand some result items may not make sense to be cached, as there may be too many of valid input parameter combinations (e.g. now there are 900M of \textit{(run, file)} combinations).

Other way to boost performance is pre-fetching most common queries from time to time (determined manually and/or automatically). Also it's possible have different validity dates for certain fields. if no volatile fields are not explicitly requested, even a very old cache could be used

%%%%%  -------------------------------------------
\subsubsection*{Data providers' performance issues}
In the cases when new records are coming but the existing ones are not changing much, continuous view maintenance that computes only differences from earlier results could be a good solution to improve the performance especially for popular and heavy queries containing joins and/or aggregations.

This is exactly the case for the most popular expensive query (see \hyperref[run_dataset_heavy]{query \#\ref{run_dataset_heavy} in \ref{appendix_das_cli_logs} DAS log analysis}) over DBS system (80GB + 280GB indexes): 'find files where run in [r1, r2, r3] and dataset=X' that needs to join all the biggest tables in that database.

\begin{verbatim}
Dataset (164K rows) -> Block (2M) -> Files (31M) -> FileRunLumi (902M) <- Runs (65K)
\end{verbatim}

Having a materialized view with all these tables joined together would allow answering such queries much quicker. Given low change rates (in comparison to data already present), maintaining the view should be also comparatively cheap with the only expense of just couple of times of storage space (storage is bound by the size of the largest table anyway).

For Oracle \textit{materialized refresh fast views with query rewriting} would be completely transparent and would not need any changes to proprietary systems, but it has a couple of limitations on the queries\cite{Oracle11}. Otherwise some other continuous view maintenance tool (e.g. DBToaster\footnote{http://www.dbtoaster.org that is being developed at EPFL}) could be used, however this wouldn't be as transparent.

%%%%%  -------------------------------------------

\subsubsection*{Integrating distributed information efficiently}
Bloom-join (which could be quite transparent and implemented even on DB side (efficient pure SQL is doable for MySQL, and probably for Oracle too) quite almost transparently (a helper function could take the initial query and bloom-filter's bit-vector as parameters).