### 2012-12-21 doc/thesis/latex/milestones.tex
elsevier styles; minor changes

\begin{appendix}
\section{Project Milestones}

Initially written up by Valentin (technical lead of the project from US):

\begin{footnotesize}
\begin{Verbatim}[commandchars=\\\{\}]
Sep:
   - familiarize with techniques, research activities in the field
   - setup all tools and learn code to do actual work
   - describe and put on paper CMS use case
   - propose several approaches to target the project
Oct:
   - pick-up couple of approaches for development
   - develop initial prototype in high-level
       - outline workflow
   - math proof of choosen approaches
   - present results at EPFL & CMS DMWM team
Nov:
   - made a choice among target approach and start working on real prototype
   - it would be nice to see math foundations in place, including 
     inital benchmarks, time estimates, stats, etc.
   - mid-term write-up document about current achievements, 
     description of choosen approach and its implementation milestones
   - at this stage work with couple of data-providers (or their limited scope) 
     to demonstrate capabilties of the chosen approach
   - I do not expect at this stage any optimizations, but would like to hear 
     where those can be done, including strategies, technical improvements, etc.
   - do not limit yourself to CMS use case, explore possibility to apply
       choosen approach outside of CMS field
     - a toy prototype with non CMS data-providers is desired to have
            - demonstrate and discuss if it can be applied elsewhere
   - present results at EPFL & CMS DMWM team
   - Peter's action: based on the work presented, make a decision about conference
     submission (see below Jan milestone for approval)
Dec:
   - expand existing work from couple of data-providers into full set of CMS  data-services
   - work on integration with existing DAS code and prepare patch for  production deployment
   - plenty of testing, including but not limited to:
     - unit tests
     - integration tests
     - end-user UI studies
           - good to have prototype and give it to end-users
           - study and observe user behavior
   - present final strategy at EPFL & CMS DMWM team, including benchmarks, stats, user UI studies
Jan:
   - work on optimization (if required)
   - integration with DAS and production deployment
   - prepare main sections of write-up document
     - target approach & math foundation
     - prototype description
     - user studies, benchmarks, stats
   - submit write-up for EPFL & CMS review
     - if it is done, Peter will need to get some approval for conference presentation
Feb:
   - final document in place describing work, achievements, results
   - make conclusion about contribution to the field
     - discuss innovation work
     - discuss engineering work
     - discuss CMS use case and achievements
     - discuss possible application outside the field
Mar 15:
    - thesis delivery
Apr:
   - defense
\end{Verbatim}
\end{footnotesize}

\end{appendix}